# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

1. [Download the data from this location]
(https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)
2. Unzip the file and move the csv to the directory to be used as working directory.
3. Set the working directory in R with setwd()
4. read the data into a data frame:

```{r}
DF<-read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

1. Histogram of steps groupped by date (omitting NAs):

```{r}
x<-aggregate(steps ~ date, DF, FUN=sum)
hist(x$steps)
```

2. Mean and median of steps per day

```{r}
meanSteps<-mean(x$steps)
medianSteps<-median(x$steps)
```

The mean steps per day is `r as.integer(meanSteps)`

The median steps per day is `r as.integer(medianSteps)`

## What is the average daily activity pattern?

1. Plot average steps taken by interval 

```{r}
x1<-aggregate(steps ~ interval, DF, FUN=mean)
plot(x1$interval,x1$steps,type="l")
```

2. Which 5-minute interval contains the maximum number of steps

```{r}
maxSteps<-x1$interval[which.max(x1$steps)]
```

The maximum number of steps(on average) is at interval `r maxSteps`

## Imputing missing values

There are a number of missing values (NA) in the original data, in particular
these NAs affect steps:
```{r}
naSteps<-sum(is.na(DF$steps))
naDate<-sum(is.na(DF$date))
naInterval<-sum(is.na(DF$interval))
```

The number of missing entries in steps is `r naSteps`.

The number of missing entries in dates is `r naDate`.

The number of missing entries in interval is `r naInterval`.

### Imputation strategy: use average steps to populate missing values

```{r}
indx <- which(is.na(DF[, 1])==TRUE) 
DF[indx,1]<-mean(x1$steps[x1$interval==DF[indx,3]],na.rm=TRUE)
```

Histogram of steps groupped by date (After imputing):

```{r}
x<-aggregate(steps ~ date, DF, FUN=sum)
hist(x$steps)
```

Mean and median of steps per day

```{r}
meanSteps2<-mean(x$steps)
medianSteps2<-median(x$steps)
```

The mean steps per day is `r as.integer(meanSteps2)`

The median steps per day is `r as.integer(medianSteps2)`

## Are there differences in activity patterns between weekdays and weekends?

1. new dataframe with weekday included:

```{r}
DF2<-cbind(DF,weekdays(as.Date(DF$date)))
names(DF2)[4]<-"DayofWeek"
```

2. Separate weekdays and weekends

```{r}
DFwd<-DF2[which(DF2$DayofWeek!="Sunday" & DF2$DayofWeek!="Saturday"),]
DFwe<-DF2[which(DF2$DayofWeek=="Sunday" | DF2$DayofWeek=="Saturday"),]
```

3. plot

```{r}
plot(DFwd$interval,DFwd$steps,type="l")
plot(DFwe$interval,DFwe$steps,type="l")
```
